---
title: "Animate Control Chart"
slug: "annimate-control-chart"
author: "Jameson Marriott"
date: "2019-03-09"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "png"
draft: TRUE
tags:
  - r
  - SPC
---

Recently I wrote a post about [creating control charts in r](https://jmarriott.com/posts/testing-ggqc/), and now I want to experiment with animating one of those charts.

Lets start with the `tidyverse`, `gganimate`, and `ggQC`. 

```{r startup, message=FALSE, warning=FALSE}
library(tidyverse)
library(gganimate)
library(ggQC)
```

Now, lets rebuild the last control chart from my previous post. 

```{r starting_plot}
# Generate sample data
set.seed(20190117)
example_df <- data_frame(values = rnorm(n=30*5, mean = 25, sd = .005),
                         subgroup = rep(1:30, 5),
                         n = rep(1:5, each = 30)) %>%
  add_row(values = rnorm(n=2*5, mean = 25 + .006, sd = .005),
          subgroup = rep(31:32, 5),
          n = rep(1:5, each = 2)) 

violations <- example_df %>%
  QC_Violations(value = "values", grouping = "subgroup", method = "xBar.rBar") %>%
  filter(Violation == TRUE) %>%
  select(data, Index) %>%
  unique()

ggQC_example <- example_df %>%
  ggplot(aes(x = subgroup, y = values, group = 1)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  stat_QC(method = "xBar.rBar", auto.label = TRUE, label.digits = 4) +
  scale_x_continuous(expand =  expand_scale(mult = c(.05, .15))) + # Pad the x-axis for the labels
  ylab("x-bar") +
  theme_bw() +
  geom_point(data = violations, aes(x = Index, y = data), color = "red")
ggQC_example
```

The way that I overlaid the red points on top of the black points may cause us some grief when we go to annimate this plot, but let's give it a try.

```{r take_one}
ggQC_example +
  transition_reveal(subgroup)
```

Ok, setting the red points asside, the graph looks crazy because it seems to recalculate the control limits for each frame. Let's try making the same plot from a tidy dataset, without the colored points, and see if that annimates a little more nicely. 

```{r take_two, message=FALSE, warning=FALSE}
tidy_df <- example_df %>%
  group_by(subgroup) %>%
  summarise(point = mean(values)) %>%
  mutate(UCL = xBar_rBar_UCL(data = example_df, value = "values", grouping = "subgroup"),
         LCL = xBar_rBar_LCL(data = example_df, value = "values", grouping = "subgroup"))

take_two <- tidy_df %>%
  ggplot(aes(x = subgroup, y = point)) +
  geom_point() +
  geom_line() +
  geom_line(aes(x = subgroup, y = UCL), color = "red") +
  geom_line(aes(x = subgroup, y = LCL), color = "red") +
  geom_line(aes(x = subgroup, y = (UCL + LCL) / 2), alpha  = .5) +
  xlab("x-bar") +
  ylab("subgroup") +
  theme_bw() +
  transition_reveal(subgroup)
take_two
```

That worked much more nicely. If I wanted the UCL and LCL to show up early, I could make those `geom_hline`s instead. 