---
title: "Testing out ggQC"
slug: "testing-ggQC"
author: "Jameson Marriott"
date: "2019-02-11"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "svg"
draft: FALSE
---

`r` is a great language for statistical process control largely because there are several packages which make it quite easy, including [qcc](https://cran.r-project.org/package=qcc), [SixSigma](https://cran.r-project.org/package=SixSigma), and a new one that I want to explore in this post called [ggQC](http://rcontrolcharts.com/). I've been using `qcc` for a while and I trust its calculations, so I'll use that to validate `ggQC`. 

`ggQC` was created by Keneth Grey and I came accross it when I read his [excelent article](https://r-bar.net/xmr-limits-moving-range/) on why we use r-bar and a constant to calculate the control limits on a control chart. The advantage of `ggQC` is that it extends [ggplot2](https://ggplot2.tidyverse.org/), so that you get all of the flexibility that comes with that. 

First, let's load the `tidyverse`, `qcc`, and `ggQC` (you'll need to install these first if you haven't already). 

```{r startup, warning=FALSE, message=FALSE}
library(tidyverse)
library(qcc)
library(ggQC)
```

Now, let's create an x-bar chart with qcc to use as a baseline. 

```{r qcc_x-bar, message=FALSE, results='hide'}
# Generate sample data
set.seed(20190117)
example_df <- data_frame(values = rnorm(n=30*5, mean = 25, sd = .005),
                         subgroup = rep(1:30, 5),
                         n = rep(1:5, each = 30)) %>%
  add_row(values = rnorm(n=2*5, mean = 25 + .006, sd = .005),
          subgroup = rep(31:32, 5),
          n = rep(1:5, each = 2))

# Spread the data and plot the control chart
qcc_chart <- example_df %>%
  spread(key = n, value = values) %>%
  select(-subgroup) %>%
  qcc(type = "xbar", data.name = "Example Data")
qcc_chart
```

You can see that as long as the data is formatted with each group on its own row, this package makes it very easy to generate a funcitonal control chart. However, customizing this chart isn't as easy for those of us who primarily use `ggplot2` instead of base plots. 

Let's take a look at the same thing with ggQC. First we'll need to put the data in long format, then we'll pass that to `ggplot2` and add some `ggQC` layers. 

```{r ggQC_chart}
ggQC_example <- example_df %>%
  ggplot(aes(x = subgroup, y = values, group = 1)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  stat_QC(method = "xBar.rBar", auto.label = TRUE, label.digits = 4) +
  scale_x_continuous(expand =  expand_scale(mult = c(.05, .15))) + # Pad the x-axis for the labels
  ylab("x-bar") +
  theme_bw()
ggQC_example
```

You can see that both packages compute the UCL and LCL identically. 

I would like for the two stat summaries to be plotted as part of `stat_QC`, but I like that you get more control over them when you plot them on their own and that you see exacly what you're plotting. While building the chart this way seems a little tedious, it makes it easy to read what's happening and also easy to add additional layers, like overlaying the individual points:

```{r ggQC_add_individuals}
ggQC_example +
  geom_point(alpha = .25)
```

I like showing the plot like this because it is more clear to an innocent bystander exactly what is being plotted. 

One thing that I would like for ggQC to provide is an easy way to make a decent chart in just a few lines of code. For example, something like this:

```{r wish_list, eval=FALSE}
tidy_df %>%
  ggplot(aes(x = Groups, y = Observations)) +
  stat_quick_QC(method = "xBar.rBar", auto.label = TRUE, label.digits = 4)
```

Another thing missing from the ggQC package is coloring the violations. It is supposed to color them with the `stat_qc_violations()` layer, but I think that is a messy solution with the four facets. It is also buggy, as you can see by the two violations that it identifies being plotted in the wrong place on the x axis. 

```{r ggQC_violations}
ggQC_example +
  stat_qc_violations(method = "xBar.rBar")
```

So, let's try a work-around to color the violations. First we'll get the violations from the ggQC `QC_Violations` function and then we'll add them to the `ggQC_example` plot, colored red. 

```{r ggQC_violation_workaround}
violations <- example_df %>%
  QC_Violations(value = "values", grouping = "subgroup", method = "xBar.rBar") %>%
  filter(Violation == TRUE) %>%
  select(data, Index) %>%
  unique()

ggQC_example +
  geom_point(data = violations, aes(x = Index, y = data), color = "red")
```


You can see that `ggQC` doesn't use the same violation rules as `qcc`. `ggQC` lists its rules in the help documentation for `QC_Violations()`, but I don't see the rules documented anywhere for the `qcc` package. There are multiple sents of rules that people use, but I like to refer to the NIST Engineering Statistics Handbook [Western Electric rules](https://www.itl.nist.gov/div898/handbook/pmc/section3/pmc32.htm#WECO%20rules) since they're published by NIST and are publically available. However, neither package uses these rules exactly, so lets do the same work-around with the qcc package for fun. 

```{r qcc_violation_workaround}
violations <- data_frame(Observations = qcc_chart$violations) %>%
  map(unlist) %>%
  as_data_frame() %>%
  left_join(as_data_frame(qcc_chart$statistics) %>% 
              mutate(Observations = row_number()), 
            by = "Observations")

ggQC_example +
  geom_point(data = violations, aes(x = Observations, y = value), color = "red")
```

Overall I like the ggQC package. It's still an early version and has some things to work out, but I think that it's the natural progression of control charting in `r`.

This post was generated with version `r packageVersion("ggQC")` of `ggQC`.