---
title: "How to Animate a Control Chart"
slug: "annimate-control-chart"
author: "Jameson Marriott"
date: "2019-04-02"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "png"
draft: TRUE
tags:
  - r
  - SPC
---

Recently I wrote a post about [creating control charts in r](https://jmarriott.com/posts/testing-ggqc/), and now I want to experiment with animating one of those charts.

Lets start with the `tidyverse`, `gganimate`, and `ggQC`. 

```{r startup, message=FALSE, warning=FALSE}
library(tidyverse)
library(gganimate)
library(ggQC)
```

Now, lets rebuild the last control chart from my previous post. 

```{r starting_plot}
# Generate sample data
set.seed(20190117)
example_df <- data_frame(values = rnorm(n=30*5, mean = 25, sd = .005),
                         subgroup = rep(1:30, 5),
                         n = rep(1:5, each = 30)) %>%
  add_row(values = rnorm(n=2*5, mean = 25 + .006, sd = .005),
          subgroup = rep(31:32, 5),
          n = rep(1:5, each = 2)) 

violations <- example_df %>%
  QC_Violations(value = "values", grouping = "subgroup", method = "xBar.rBar") %>%
  filter(Violation == TRUE) %>%
  select(data, Index) %>%
  unique()

ggQC_example <- example_df %>%
  ggplot(aes(x = subgroup, y = values)) +
  stat_summary(fun.y = mean, geom = "line", aes(group = 1)) +
  stat_summary(fun.y = mean, geom = "point", aes(group = 1)) +
  stat_QC(method = "xBar.rBar", auto.label = TRUE, label.digits = 4) +
  scale_x_continuous(expand =  expand_scale(mult = c(.05, .15))) + # Pad the x-axis for the labels
  ylab("x-bar") +
  theme_bw() +
  geom_point(data = violations, aes(x = Index, y = data), color = "red", group = 1)
ggQC_example
```

You can see that I added some grouping variables to try and control the annimation, but I couldn't achieve the annimation that I was looking for. Take a look: 

```{r take_one, message=FALSE, warning=FALSE}
ggQC_example +
  transition_reveal(subgroup)
```

Ok, setting the red points asside, the graph looks crazy because it seems to recalculate the control limits for each frame. Let's try making the same plot from a tidy dataset, and see if that annimates a little more nicely. 

```{r take_two, message=FALSE, warning=FALSE}
# is there a way to hold the annimate for a second before it restarts?

UCL <- xBar_rBar_UCL(data = example_df, value = "values", grouping = "subgroup")
LCL <- xBar_rBar_LCL(data = example_df, value = "values", grouping = "subgroup")

take_two_df <- example_df %>%
  group_by(subgroup) %>%
  summarise(point = mean(values)) %>% # caluculate the x-bar points
  left_join(violations, by = c("subgroup" = "Index"), suffix = c("normal", "violation")) %>% # add the information about which points are out-of-control
  mutate(control = case_when( # supply missing information from the join
    is.na(data) ~ "out", 
    TRUE ~ "in"
  ))

take_two <- take_two_df %>%
  ggplot(aes(x = subgroup, y = point)) +
  geom_line() +
  geom_point(aes(group = subgroup, color = control)) + # adding the group aes keeps the points from dissapearing
  geom_hline(yintercept = c(LCL, UCL), color = "red") + # usin geom_hline keeps the line from annimating
  geom_hline(yintercept = (UCL + LCL) / 2, alpha  = .5) +
  xlab("x-bar") +
  ylab("subgroup") +
  scale_color_manual(values = c("red", "black")) + # manually color the in-control vs out-of-control points
  ggtitle("Annimated Control Chart") +
  theme_bw() +
  theme(legend.position = "none") + # the legend is self-explanatory
  transition_reveal(subgroup) # specify the annimation
  
animate(take_two, end_pause = 10) # using the animate() function allows me to set the end_pause
```

Sure enough, that worked much more nicely. The lesson-learned is that items that are calculated at the time of plotting act differently from those that are simply data being thrown on a graph. I couldn't figure out how to get `stat_QC` (which is calculated at run-time when plotting) to remain static, so I had to create a tidy dataset with all the information that I needed.

After watching this annimation run for a while, I realized that I would like it to pause on the last frame before starting over. Fortunately, I was able to find [this post](https://stackoverflow.com/questions/53092216/any-way-to-pause-at-specific-frames-time-points-with-transition-reveal-in-gganim) on stack overflow which explains how to accomplish that. Basically, you have to add a bunch of duplicate rows at the end so that the animation doesn't look like it's doing anything. This is a waste of frames, but is a quick fix. 

```{r}
take_two_df %>%
  mutate(show_time = case_when(subgroup %in% n() ~ 10, # add the number 10 to the last row
                               TRUE ~ 1)) %>% # add 1 to the rest of the rows
  uncount(show_time) %>% # dupicate the rows show_time number of times
  mutate(reveal_time = row_number()) %>%
  ggplot(aes(x = subgroup, y = point)) +
  geom_line() +
  geom_point(aes(group = subgroup, color = control)) + # adding the group aes keeps the points from dissapearing
  geom_hline(yintercept = c(LCL, UCL), color = "red") + # usin geom_hline keeps the line from annimating
  geom_hline(yintercept = (UCL + LCL) / 2, alpha  = .5) +
  xlab("x-bar") +
  ylab("subgroup") +
  scale_color_manual(values = c("red", "black")) + # manually color the in-control vs out-of-control points
  ggtitle("Annimated Control Chart") +
  theme_bw() +
  theme(legend.position = "none") + # the legend is self-explanatory
  transition_reveal(reveal_time) # use the new reveal_time column to control the animation
```


Regardless, `ggannimate` makes annimations way easier than trying to creating all of those individual plots on their own and then combining them into a .gif. It would only be 32 plots if I wanted it to jump from point-to-point, but notice that the lines also get drawn rather than jumping. 