---
title: "XKCD-Gutenbergr Passwords"
slug: "Passwords"
author: "Jameson Marriott"
date: "2020-11-16"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "svg"
draft: TRUE
tags:
  - R
---

I have been inspired by [this informative XKCD comic](https://xkcd.com/936/) on password security for a few years and I often use its advice by using random-word-generating websites. But, I have to wonder what dictionaries these sites use and how random the words are that they spit out. So, I decided to write a bit of code so that I could control the process. I also thought that it would be fun to use books from [https://www.gutenberg.org/](https://www.gutenberg.org/) as the word sources. As long as I chose books somewhat randomly then I think that should be pretty secure. 

First I will load the required packages. 

```{r setup}
library(tidyverse)
library(gutenbergr)
library(tidytext)
```

Since [Pride and Prejudice](https://www.gutenberg.org/ebooks/1342) is currently the most downloaded book on gutenberg over the past 30 days and I happen to like it myself, I'll use that as my example book. It is easy to download the full text with the `gutenbergr` package. 

```{r}
test <- gutenberg_download(1342)

data("stop_words") # load stop words

test %>%
  unnest_tokens(word, text) %>%
  mutate(word = str_extract(string = word, pattern = "[[:alpha:]]+")) %>% #remove any non-alphanumeric characters. 
  anti_join(stop_words) %>%
  select(word) %>%
  unique() %>%
  drop_na()
  
```





















