---
title: "XKCD-Gutenberg Passwords"
slug: "xkcd-passwords"
author: "Jameson Marriott"
date: "2020-12-20T20:38:00-06:00"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "svg"
draft: FALSE
tags:
  - R
---

I have been inspired by [this informative XKCD comic](https://xkcd.com/936/) on password security, and I often follow its advice by using random-word-generating websites. But I have to wonder what dictionaries these sites use and how random the words are that they spit out. So I thought that it would be fun to make my own generator using books from [https://www.gutenberg.org/](https://www.gutenberg.org/) as my dictionaries, that way I would at least know where they are coming from. As long as I chose books somewhat randomly, I think that should be pretty secure. 

First I will load the required packages. 

```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(gutenbergr)
library(tidytext)
```

Since [Pride and Prejudice](https://www.gutenberg.org/ebooks/1342) is currently the most downloaded book on gutenberg over the past 30 days and I happen to like it myself, I'll use that as my example book. It is easy to download the full text with the `gutenbergr` package. 

```{r password, message=FALSE, warning=FALSE}
book_text <- gutenberg_download(1342, mirror = "http://mirrors.xmission.com/gutenberg/")

data("stop_words") # load stop words

book_text %>%
  unnest_tokens(word, text) %>% # turn the text into a single column of words
  mutate(word = str_extract(string = word, pattern = "[[:alpha:]]+")) %>% # remove any non-alphanumeric characters 
  select(word) %>% # get rid of the extra columns
  unique() %>% # get rid of duplicate words
  anti_join(stop_words, by = "word") %>% # get rid of boring "stop" words
  drop_na() %>% # drop anything that didn't make it through cleanly
  unlist() %>% # turn the column into a vector that sample() knows what to do with
  sample(4) # chose four words at random
```

Well that was super easy, wasn't it? The only thing that isn't easy with this setup is looking up a book to use. But that can readily be solved with a shiny app. You can see the code below (at the time this post was knit---the current code will always be on github [here](https://github.com/jameson-marriott/jmarriott.com/blob/master/shiny/password_app.R)), and the live app is at [https://jameson-marriott.shinyapps.io/Password_App/](https://jameson-marriott.shinyapps.io/Password_App/).

```{r shiny, code=readLines("~/R/jmarriott.com/shiny/password_app.R"), eval=FALSE}

```

















