---
title: "Evenly Spaced Month Charts"
slug: "evenly-spaced-month-charts"
author: "Jameson Marriott"
date: "2024-04-25"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "svg"
draft: TRUE
tags:
  - R
---

I recently noticed that [ggplot2](https:://ggplot2.tidyverse.org) spaces date axes literally even when grouped by month. I've been using [ggplot2](https:://ggplot2.tidyverse.org) extensively for years and I don't remember noticing before, so this is not really a big deal, but now that I know it bugs me a lot. Take a look below. 

As always, let's load the `tidyverse`.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
```

Next, we need some example data. Nothing fancy, just something to plot as an example. 

```{r data}
month_data <- tibble(dates = seq(ymd("2023-11-01"), ymd("2024-04-30"), by = "day"),
                     data = sample(1:1e3, 182)) |>
  mutate(month = floor_date(dates, "month")) |>
  group_by(month) |>
  summarise(monthly_mean = mean(data))

month_data
```

Now let me show you how the distance between the months is not equal. 

```{r plot}
month_data |>
  ggplot(aes(x = month, y = monthly_mean)) +
  geom_col(width = 29) +
  ggtitle("The Months Are Not Spaced Evenly!") +
  # add the year to the first month plotted in each year
  scale_x_date(labels = scales::label_date_short())
```

How frustrating! The fix is not that hard, though. We are going to change the date axis to factors with a neat little trick. We're going to hack the `label_date_short()` function to change the dates to text and then change the dates to factors. Here we go. 

```{r fix_the_data}
# get the label_date_short() function
prettify_dates <- scales::label_date_short()

# we use the new function we made and make the dates factors so that they sort 
# correctly. 
new_month_data <- month_data |>
  mutate(pretty_month = prettify_dates(month),
         pretty_month = as_factor(pretty_month))

new_month_data
```

Now all we have to do is drop the date scale and let `geom_col()` chose its own width. 

```{r better_plot}
new_month_data |>
  ggplot(aes(x = pretty_month, y = monthly_mean)) +
  geom_col() +
  ggtitle("The Spacing Is Even Now!")
```

