---
title: "Modeling Operating Characteristic Curves in R"
slug: "oc-curves"
author: "Jameson Marriott"
date: "2019-09-13T06:37:00-06:00"
output:
  blogdown::html_page:
    toc: true
    fig_width: 6
    dev: "png"
draft: TRUE
tags:
  - R
  - Data
  - Sampling
---



```{r startup, message=FALSE, warning=FALSE}
library(tidyverse)
library(AcceptanceSampling)
```

Before we setup a simulation, let's us the `AcceptanceSampling` package to generate an OC curve for us. 

```{r oc_curve}
oc_obj <- AcceptanceSampling::OC2c(n = 13, c = 0)
plot(oc_obj, xlim = c(0, .5))
```

Notice that with c = 0, which is becoming the standard in inspection sampling plans, the OC curve doesn't have the usual backwards "S" shape, but instead takes an imidiate dive down from the 100% probability of acceptance.

The first thing that I want to do is simulate the traditional OC curve that was so easily provided by the `AcceptanceSampling` package. 

```{r}
sim1 <- tibble(`Percent Defective` = rep(seq(0, .5, .01), each = 100000)) %>%
  mutate(`n = 05`  = map(`Percent Defective`, rbinom, n = 5,  size = 1), # prob = `Percent Defective`
         `n = 10` = map(`Percent Defective`, rbinom, n = 10, size = 1), # prob = `Percent Defective`
         `n = 15` = map(`Percent Defective`, rbinom, n = 15, size = 1), # prob = `Percent Defective`
         `n = 20` = map(`Percent Defective`, rbinom, n = 20, size = 1)) %>% # prob = `Percent Defective`
  gather(key = n, value = Sample, - `Percent Defective`) %>%
  mutate(`Probability of Acceptance` = map_int(Sample, max)) %>%
  group_by(`Percent Defective`, n) %>%
  summarise(`Mean Probability of Acceptance` = 1 - mean(`Probability of Acceptance`))

sim1 %>%
  ggplot(aes(x = `Percent Defective`, y = `Mean Probability of Acceptance`, color = n)) +
  geom_point(alpha = .5) +
  geom_line() +
  labs(title = "Simulated OC Curve", 
       subtitle = "c = 0",
       caption = "jmarriott.com") +
  ylab("Probability of Acceptance") +
  xlab("Percent Defective") +
  theme_bw()
```

Using `each = 100000` makes the graph look nice and smooth, but otherwise doesn't seem to be much of an improvement over `each = 10000` or even `each = 1000` and it takes a lot longer to run. 